---
title: "Mechanism Model Data Prep"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries
```{r}
library(tidyverse)
library(janitor)
library(caret)
library(tidymodels)
```

Read in data
```{r}
to_comb_cardio_syn = c("Cardiovascular event", "Other syncope or dyspnea")
to_comb_hit_obj = c("Hit object","Hit player","Hit with ball","Hit with racquet/paddle")

neiss_df_fct_d = read_rds("../clean_data/clean_full.RDS")
```


Custom Functions
```{r}
preprocess_neiss <- function(neiss_df, top_diags, top_parts, cols_to_use, is_new = FALSE) {
  
  return_df = NULL
  
  if(!is_new){
    cols_to_use = c(cols_to_use, "primary_mechanism")
  }
  
  return_df =  neiss_df |>  
    mutate( body_part = as.character(body_part)
         , diagnosis = as.character(diagnosis)
         , body_part = ifelse(body_part %in% top_parts, body_part, "other")
         , diagnosis = ifelse(diagnosis %in% top_diags, diagnosis, "other")
         , narrative = tolower(narrative)
         , sex = factor(sex)
         , body_part = factor(body_part)
         , diagnosis = factor(diagnosis)
         ) |>
    select(all_of(cols_to_use))
  
  return(return_df)
}

get_keywords <- function(neiss_df, word_funct = str_detect ) {
  
  fall_words <- "slip|trip|fall|dive|dove|fell|d&|foosh"
  hit_rack_words <- "hit|paddle|racket|raquet|bat"
  hit_ball_words <- "ball|hit with ball"
  five_words <- "musculo|contusion|epist"
  mov_words <- "lunge|bend|hyperex|disloc|sprain|strain|bump|tear|pop|snap|roll|cut|abrasion|sting|jam"
  heat_words <- "heat|hot|dehydrat"
  faint_words <- "syncope|dizz|dyspnea" # INTERACT WITH HEAT WORDS?
  cardio_words <- "heart|rhythm|angina|chest|chest pain|dvt|atax|weak|blood pressure|palpitat"
  hit_other <- "fence|wall|tree|net|bench|ran into|ran"
  hit_player <- "ran into|another|player|collieded"
  other_sport <- "been playing|volly|baseball|basketball|soccer|football|golf|sports"

return_df = NULL

return_df <- neiss_df %>% 
  mutate(
    twist = as.integer(word_funct(narrative, "twist")),
    fall_words  = as.integer(word_funct(narrative, fall_words)),
    hit_rack_words = as.integer(word_funct(narrative, hit_rack_words)),
    hit_ball_words = as.integer(word_funct(narrative, hit_ball_words)),
    five_words = as.integer(word_funct(narrative, five_words)),
    mov_words = as.integer(word_funct(narrative, mov_words)),
    heat_words = as.integer(word_funct(narrative, heat_words)),
    faint_words = as.integer(word_funct(narrative, faint_words)),
    cardio_words = as.integer(word_funct(narrative, cardio_words)),
    hit_other = as.integer(word_funct(narrative, hit_other)),
    hit_player= as.integer(word_funct(narrative, hit_player)),
    other_sport = as.integer(word_funct(narrative, other_sport))
  ) |>
  select(-narrative)
  
return(return_df)
}
```


Filter for new data, using previously labeled data minus fatalities
```{r}
pb_model_df = neiss_df_fct_d |>
  filter(id == "pickleball" & disposition != "FATALITY INCL. DOA, DIED IN ER")# & age >= 50) #|> count(primary_mechanism)

tennis_model_df <- neiss_df_fct_d |>
  filter(id == "tennis" & disposition != "FATALITY INCL. DOA, DIED IN ER") 


consolidated_model_df = pb_model_df |> 
  bind_rows(tennis_model_df) |> 
  mutate(
    primary_mechanism = ifelse(primary_mechanism %in% to_comb_cardio_syn, "cardio/sync", primary_mechanism)
    ,primary_mechanism = ifelse(primary_mechanism %in% to_comb_hit_obj, "Hit with various obj", primary_mechanism)
         , primary_mechanism = factor(primary_mechanism)     
         )  |>
  select(cpsc_case_number, psu , stratum , weight , treatment_date
         ,primary_mechanism
         ,age, sex ,body_part, diagnosis
#         , body_region, diagnosis_2, other_diagnosis # Possibly add later
         ,narrative
         )

#consolidated_model_df

# tmp_r = consolidated_model_df |> apply(2, \(x) sum(is.na(x)) )
# 
# tmp_r
# tmp_r / nrow(consolidated_model_df)
# 
# tmp_n = consolidated_model_df |> apply(2, n_distinct )
# 
# tmp_n
# tmp_n / nrow(consolidated_model_df)


#consolidated_model_df |> count(other_diagnosis, sort = T) |> View()
```


Get new data, using the same filters used to get the original data
```{r}
clean_data <- read_rds("C:/Users/Jake/Desktop/Exercise_Equipment/clean_data/clean_neiss_df_2013-2022.RDS") |> mutate(age = as.numeric(age_in_years) ) |>
  filter(Treatment_Date >= ymd("2020-01-01") & Disposition != "FATALITY INCL. DOA, DIED IN ER")

tennis_dat = clean_data  |>
  filter(Product_1 == "TENNIS (ACTIVITY, APPAREL OR EQUIPMENT)" | Product_2 == "TENNIS (ACTIVITY, APPAREL OR EQUIPMENT)" | Product_3 == "TENNIS (ACTIVITY, APPAREL OR EQUIPMENT)") |> 
  filter(age >= 60) |> mutate(id = "tennis")

pball_words = "pickleball|pickle ball|pickelball|pickel ball"

pball_dat = clean_data  |>
  filter(str_detect(string = tolower(Narrative), pattern = pball_words  )) |>  
  mutate(id = "pickleball")

new_data_clean = tennis_dat |>
  bind_rows(pball_dat) |> janitor::clean_names() |> mutate(age = as.integer(age)) 

```


Get used column names and used body parts and diagnosis
```{r}
#consolidated_model_df
m_names = select(consolidated_model_df,-primary_mechanism) |> names()

new_data_clean |> select(all_of(m_names))

parts_use = consolidated_model_df |> count(body_part) |> filter(n > 5) |> pull(body_part)
diag_use =  consolidated_model_df |> count(diagnosis) |> filter(n > 5) |> pull(diagnosis)
model_cols = select(consolidated_model_df,-primary_mechanism) |> names()
```


Use custom functions to preprocess both new and old data
```{r}
model_df_raw = preprocess_neiss(consolidated_model_df, top_parts = parts_use, top_diags = diag_use, cols_to_use = model_cols )

new_df_prepro = preprocess_neiss(new_data_clean, top_parts = parts_use, top_diags = diag_use, cols_to_use = model_cols, is_new = TRUE )
```



```{r}
model_df_final_detect = model_df_raw |> get_keywords()
model_df_final_count = model_df_raw |> get_keywords(word_funct = str_count)
new_data_to_model_detect = new_df_prepro |> get_keywords()
new_data_to_model_count = new_df_prepro |> get_keywords(word_funct = str_count)
```



```{r}
narrative_tbl = select(model_df_raw, cpsc_case_number, narrative ) |>
  bind_rows(select(new_df_prepro, cpsc_case_number, narrative )) |>
  count(cpsc_case_number, narrative) |>
  select(-n) 



```

```{r}
narrative_tbl |>
  write_csv("../data_to_model/narrative_tbl_all.csv")
```


Export clean data for modeling
```{r}
model_df_final_detect |> write_csv("../data_to_model/model_df_final_detect.csv")
model_df_final_count |> write_csv("../data_to_model/model_df_final_count.csv")
new_data_to_model_detect |> write_csv("../data_to_model/new_data_to_model_detect.csv")
new_data_to_model_count |> write_csv("../data_to_model/new_data_to_model_count.csv")

model_df_final_detect |> write_rds("../data_to_model/model_df_final_detect.rds")
model_df_final_count |> write_rds("../data_to_model/model_df_final_count.rds")
new_data_to_model_detect |> write_rds("../data_to_model/new_data_to_model_detect.rds")
new_data_to_model_count |> write_rds("../data_to_model/new_data_to_model_count.rds")

```


Export full and narrative data for Keras text preprocessing
```{r}
#model_df_raw |> count(primary_mechanism)
#model_df_raw  |>  mutate(primary_mechanism = as.numeric(primary_mechanism)) |> count(primary_mechanism)

# model_df_raw |> # mutate(primary_mechanism = as.numeric(primary_mechanism)) |> #count(primary_mechanism)
#   write_csv("../data_to_model/keras_data/model_df_fullv2.csv")
# 
# new_df_prepro |> 
#   write_csv("../data_to_model/keras_data/new_model_df_fullv.csv")

model_df_raw |>  #mutate(primary_mechanism = as.numeric(primary_mechanism)) |> 
  select(cpsc_case_number, narrative, primary_mechanism) |>
  write_csv("../data_to_model/keras_data/model_df_full_narrative.csv")

# new_df_prepro |> 
#     select(cpsc_case_number, narrative) |>
#   write_csv("../data_to_model/keras_data/new_model_df_full_narrative.csv")

```



Training + Testing splits and binary outcome
```{r}

model_df_final_count_bin = model_df_final_count |>
  mutate(primary_mechanism_bin = ifelse(as.character(primary_mechanism) == "Multi-sport", "Multi-sport", "Other")
         ,primary_mechanism_bin = factor(primary_mechanism_bin)
         ) |>
  select(-primary_mechanism)

```

```{r}
set.seed(4)

init_split <- initial_split(model_df_final_detect, prop = .85, strata = "primary_mechanism")

train_df_detect <- training(init_split)
test_df_detect <- testing(init_split)

train_df_count <- model_df_final_count[init_split$in_id,]
test_df_count <- model_df_final_count[-init_split$in_id,]
```


```{r}
set.seed(4)

init_split_bin <- initial_split(model_df_final_count_bin, prop = .85, strata = "primary_mechanism_bin")

train_df_count_bin <- training(init_split_bin)
test_df_count_bin <- testing(init_split_bin)
```


```{r}
train_df_count |> write_csv("../data_to_model/Train and Test Sets/train_df_count.csv")
test_df_count |> write_csv("../data_to_model/Train and Test Sets/test_df_count.csv")

train_df_detect |> write_csv("../data_to_model/Train and Test Sets/train_df_detect.csv")
test_df_detect |> write_csv("../data_to_model/Train and Test Sets/test_df_detect.csv")


train_df_count_bin |> write_csv("../data_to_model/Train and Test Sets/train_df_count_bin.csv")
test_df_count_bin |> write_csv("../data_to_model/Train and Test Sets/test_df_count_bin.csv")

```



